Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-17 special-issue-ci/Response to reviewers.tex

\begin{quote}
Comments to Authors:

1) I like the state of the art which includes comprehensive overview of the most important papers related
to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
with its characteristic provided on page 4. Namely:

 - Heterogeneity - it is true that almost all devices can participate in the web browser based
 distributed computing system, providing that a user did not disable Java Script in its browser.
 However if the authors plan to use HTML5 WebWorkers only modern web browsers can
 participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or
 Windows Phone) are also excluded.

That is true, but it makes it anyway the most inclusive system so
far, as we have tested it with smartphones, tablets and computers. In
general, nowadays with rich internet applications disabling JavaScript
is not common, but anyway we do not aim for {\em all} clients, we have
enough with {\em most}. Windows Phone, for instance and maybe
fortunately, is not too common.

\begin{quote}
 - Adaptiveness - this, beside safety, is the main drawback of the system. Although it is possible to
 implement evolutionary algorithm in that way that it wil automatically adopt to the problem type
 and the environment in which it is computed, the authors do not mention about how real
 adaptiveness of the system can be achieved. The same is with scalability. Only 5 different users
 that permanently participated in the experiments are not enough to assess the scalability of the
 system, what is a crucial parameter for distributed computing.
\end{quote}

We have performed a new experiments in which we have had hundreds of
simultaneous users. In fact, with the experiments already in the
previous version of the papers 5 was a central measure, but there were
peaks of dozens. 
Adaptativity is by design, since it uses any browser and does not need
any special provision for new systems. This has also been explained in
th expanded text.

@@

We have performed a new experiments in which we have had hundreds of
simultaneous users. In fact, with the experiments already in the
previous version of the papers 5 was a central measure, but there were
peaks of dozens. 
Adaptativity is by design, since it uses any browser and does not need
any special provision for new systems. This has also been explained in
th expanded text.

\begin{quote}

 - Safety - in my opinion is the biggest concern in such kind of distributed systems. First of all the
 code of evolutionary algorithm is sent to the client as a source code in Java Script, so some
 experienced user may quite simple decode what has to be computed. It can may come into
 possession of sensitive data or/and alter this code in order to disrupt the computations or make
 them worhless. The results send to the server may also be captured and/or altered without any
 problem since the transmission is not secure.

 In my opinion you should at least highlight above issues.
\end{quote}

We have also justified our {\em pragmatic} approach to security
increasing trust by the user, but also monitoring their behavior to
detect it and eventually ban users. So far, this has not happened.

\begin{quote{
2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
maximum. In the experiments you counted how many times evolutionary algorithm was able to find
global maximum for the problem of size n=40. Initially you have found that if the population in an
evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run
(page 4, if I understood it correctly).
\end{quote}
I fact 3.46 is the average for all 50 runs. We have enphasized the
word average to make that clearer.

\begin{quote}
Thus, what was the point of using distributed computing to solve such simple problem? Why the
authors did not perform their experiments for larger values of n, what could justify distributed problem
solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
it is still behind the desktop computer in more than 60% cases (when compared to the second
baseline). I suggest to perform additional computing experiments for some higher value of n in the
"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
to focus the paper on the behavior analysis of volunteers participating in a distributed computing
system, what may be also interesting an valuable for the potential
readers.
\end{quote}

In fact, that was never the point of the paper. We have chosen Trap as
a relatively difficult problem that allows us mainly to test scaling
and measure behavior. In a production environment, a volunteer-only
environment would only be used if there is absolutely no other
resource available; in general, it would be combined with the rest of
the resources available to the experimenter: cloud or local; it would
make no sense to not use them if they are available and in fact, NodIO
allows to combine them seamlessly in a single system. That has been
explained a bit better in the introduction and conclusions
section. Besides, we have added a more difficult probmem, the shifted
Rastrigin function, to make measures in a different scenario (one in
which the problem is not solved in a few seconds).

\begin{quote}
3) Regarding the graphs I ahve already put my comments in Style section, but I have one question. In
Fig. 3 in the middle graph (for experiments conducted on 4/24/2015) the is a lot of red dots,
meaning probably that IPs from 16 to 29 did not performed any computations. What was the reason
for that?

### 2015-11-17 special-issue-ci/Response to reviewers.tex

\begin{quote}
Thus, what was the point of using distributed computing to solve such simple problem? Why the
authors did not perform their experiments for larger values of n, what could justify distributed problem
solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
it is still behind the desktop computer in more than 60% cases (when compared to the second
baseline). I suggest to perform additional computing experiments for some higher value of n in the
"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
to focus the paper on the behavior analysis of volunteers participating in a distributed computing
system, what may be also interesting an valuable for the potential
readers.
\end{quote}

In fact, that was never the point of the paper. We have chosen Trap as
a relatively difficult problem that allows us mainly to test scaling
and measure behavior. In a production environment, a volunteer-only
environment would only be used if there is absolutely no other
resource available; in general, it would be combined with the rest of
the resources available to the experimenter: cloud or local; it would
make no sense to not use them if they are available and in fact, NodIO
allows to combine them seamlessly in a single system. That has been
explained a bit better in the introduction and conclusions
section. Besides, we have added a more difficult probmem, the shifted
Rastrigin function, to make measures in a different scenario (one in
which the problem is not solved in a few seconds).

\begin{quote}
3) Regarding the graphs I ahve already put my comments in Style section, but I have one question. In
Fig. 3 in the middle graph (for experiments conducted on 4/24/2015) the is a lot of red dots,
meaning probably that IPs from 16 to 29 did not performed any computations. What was the reason
for that?
\end{quote}

The $x$ axis runs overs number of computers participating in an
experiment; that is, $4$ plots the time taken for all experiments in
which 4 computing nodes participated and the red dot is the average;
it shadows the time in the case there is a single ocassion. In the
case you mention, there are no experiments in which there were exactly
17 nodes, for instance. The ``red dots'' indicate that there was a
single experiment in which that many nodes participated. We have
clarified the caption to Figure 3 to make this clearer.

\begin{quote}
For Fig. 5 I would also avoid comparisons between the NodIO system with the SETI@home of BOINC
systems - the number of participants is completely incomparable and also the type of problems
solved by the systems are of different nature and complexity.
\end{quote}

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Commenting on the value of experiments
And also the focus. Closes #51.
Explains dots
Closes #52

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with a success rate upgraded to 100\% and an average duration of $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
characteristics: 
\begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
\caption{Duration of experiments vs. number of different IPs (nodes)
  participating in it, with averages and standard deviation shown as
  red dots. From left to right, experiments 4/4, 4/24 and 7/31. Shades
of blue indicate how many experiments included that many unique IPs.}
\label{fig:duration}
\end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
\caption{Duration of experiments vs. rank, with $y$ axis in a
  logarithmic scale. Dot color is related to the number of IPs
  participating in the experiment. From left to right, experiments
  4/4, 4/24 and 7/31.} 
\label{fig:zipf:os}
\end{figure}
%
We will have to analyze experimental data a bit further to find out why
this happens and also if there are some patterns in the three sets of
experiments. An interesting question to ask, for instance, is if
adding more computers makes the experiment take less. In fact, as
shown in Figure \ref{fig:duration}, {\em adding} more computers does
not seem to contribute to decreasing the time needed to finish the
experiment. However, the cause-effect relationship is not clear at
all. It might be the opposite: since experiments take longer to finish
and might in fact be abandoned with no one contributing for some time,
that increases the probability of someone new joining them. In fact,
with experiments taking a few seconds and due to the way the
experiments are announced, it is quite difficult that several
volunteers join in in such a short period of time, even more if we take
into account that volunteers are not {\em carried over} from previous
experiments. This implies that it would be convenient to use a problem
of a bigger size to check this hypothesis as we have done in the next
experiment. 
%  however, at this point we
% have not found this convenient since there are several other issues
% that have to be solved, as it will be shown next. %Creo que este párrafo
%saldría sobrando ahora que hicmios un experimento más grande

>>>>

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
characteristics: 
\begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-ips-OS-7-31.png}
\caption{Duration of experiments vs. number of different IPs (nodes)
  participating in it, with averages and standard deviation shown as
  red dots; in the case there is a single red dot, there was a single
  experiment in which so many computers participated (for instance, 16
  computers in the experiment in the far left or 29 in the middle
  one). 
From left to right, experiments 4/4, 4/24 and 7/31. Shades
of blue indicate how many experiments included that many unique IPs,
so lighter shade for a column of dots indicates that particular number
of computers happened less frequently, darker shadow more times.}
\label{fig:duration}
\end{figure}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-4.png}
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-4-24.png}
\includegraphics[width=0.32\linewidth]{time-vs-rank-OS-7-31.png}
\caption{Duration of experiments vs. rank, with $y$ axis in a
  logarithmic scale. Dot color is related to the number of IPs
  participating in the experiment. From left to right, experiments
  4/4, 4/24 and 7/31.} 
\label{fig:zipf:os}
\end{figure}
%
We will have to analyze experimental data a bit further to find out why
this happens and also if there are some patterns in the three sets of
experiments. An interesting question to ask, for instance, is if
adding more computers makes the experiment take less. In fact, as
shown in Figure \ref{fig:duration}, {\em adding} more computers does
not seem to contribute to decreasing the time needed to finish the
experiment. However, the cause-effect relationship is not clear at
all. It might be the opposite: since experiments take longer to finish
and might in fact be abandoned with no one contributing for some time,
that increases the probability of someone new joining them. In fact,
with experiments taking a few seconds and due to the way the
experiments are announced, it is quite difficult that several
volunteers join in in such a short period of time, even more if we take
into account that volunteers are not {\em carried over} from previous
experiments. This implies that it would be convenient to use a problem
of a bigger size to check this hypothesis as we have done in the next
experiment. 
%  however, at this point we
% have not found this convenient since there are several other issues
% that have to be solved, as it will be shown next. %Creo que este párrafo
%saldría sobrando ahora que hicmios un experimento más grande
